### 2015-10-16 Paper_v2.tex
Robustness Updates

@@

The results of the model suggest that an inhibition of either the PI3K pathway or the MAPK pathway in combination with an AR inhibitor as a possible therapy for CRPC. 
Sensitivity analysis revealed no change in the top sensitive species in the presence or absence of the AR inhibitor, enzalutamide. 
PI3K/AKT and MAPK species continued to fall in the top two percent of sensitive species. 
A study by Carver \textit{et al.} looked at dual inhibition of AR and PI3K signaling in LNCaP cells and in a Pten-deficient murine prostate cancer model \cite{Parker2013}. 
Using both the PI3K inhibitor, BEZ235, and the AR inhibitor, MDV3100 (enzalutamide), the group saw a drastic decrease in the total number of cells. 
Each inhibitor on it’s own had a much smaller effect on total cell number. They saw an increase in the cell death marker, c-PARP, in the dual inhibition case. 
The group hypothesized that AKT inhibition leads to increased AR signaling activity through increased protein concentrations of HER3. 
On the other hand AR inhibition leads to increased AKT activity due to the down regulation of PHLPP, a protein phosphatase that regulates AKT. 
For the simplicity of this model, the HER3 pathway and also cell death were not included in the model. 
Dual knock-out studies of PI3K and AR in our model show only a slight additive effect on cell cycle protein cyclin D through the dual knock-out compared to solely inhibiting PI3K (Fig. \ref{fg:Dual_Knockout}).
The triple case, with PI3K, AR, and RAF knock-outs, also only showed a slight additive effect (Fig. \ref{fg:Triple_KO}).
%The pathways in our model appear to be uncoupled and therefore little synergy is shown in the dual knock-out case. 
This could indicate that the combined decrease in cell population due to the dual inhibition of PI3K and AR is entirely due to cell death. 
The Carver \textit{et al.} study did not consider cell cycle proteins or cell growth. 
Our model does show a decrease in cell cycle proteins in the PI3K knock-out as well as in the PI3K and AR dual knock-out case in some ensemble members (Fig. \ref{fg:Dual_Knockout}). 
This result seems to be consistent with the decreased cell count in the PI3K knock-out case which is not dependent on cell death, as c-PARP levels are low. 
The decrease in cell cycle proteins in the model is due to a decrease in general translation, including free eIF4E levels and activated 40S ribosome subunit. 
The decrease in p70 (S6) activation due to inhibition of PI3K is shown in both the model and in the Carver \textit{et al.} study , indicating this result is due to the PI3K pathway (Fig. \ref{fg:Supp_KO}). 
Our experimental results confirm the Carver \textit{et al.} study in that a dual inhibition of AR and PI3K signaling led to a more prominent decrease in cell viability then each of the inhibitors alone in LNCaP cells. 
We extended the study to look at the addition of a third inhibitor, sorafenib, that inhibits Raf kinase in the MAPK pathway and an additional cell line, C4-2, which was CR.  In both cell lines, there was not a significant decrease in cell viability between the three dual inhibitor cases and the triple inhibition case at 74 hours, indicating that a dual inhibition (PI3K/AR, AR/MAPK, or PI3K/MAPK) may be a sufficient treatment.

@@

\begin{figure}\centering
\includegraphics[width=1.0\textwidth]{./figs/Figure7_ExperimentalResults.pdf}
\caption{Experimental results for multiple drug combinations on two prostate cancer cell lines, LNCaP and C4-2. A. Western blot analysis of AR, pS6, AKT, pAKT, ERK and cleaved PARP in LNCaP and C4-2 cell lines treated for 24 hrs with DMSO (control), sorafenib (5 $\mu$M), LY294002 (40 $\mu$M), and MDV3100 (10 $\mu$M) alone or in combination (at least 3 repeats). B. Cells (LNCaP and C4-2) were treated for 24, 48 and 72 hrs with sorafenib (5 $\mu$M), LY294002 (40 $\mu$M), and MDV3100 (10 $\mu$M) and cell viability was measured using MTT Assay. Values were normalized to DMSO (control). C. Cell viability results for LNCaP and C4-2 cells at varying concentration of sorafenib, LY294002, and MDV3100 after 24 hrs of treatment. Values were normalized to DMSO (control). Error bars represent standard error (at least 3 repeats with triplicates performed in each experiment).}
\label{fg:Experiment}
\end{figure}

@@

\begin{figure}\centering
\includegraphics[width=1.0\textwidth]{./figs/Fig_8_BoxPlot_KnockOut_Cases_new.pdf}
\caption{Robustness analysis of a population of CR prostate models with seven protein knock-out cases (N = 500). A log fold change of greater then zero implies that the concentration of the protein increased with the knock-out, while a log fold change of less then zero indicates that the concentration of protein decreased. A log of fold change equal to 0, shows no response due to the knock-out.
 A.,B.,C.,D. Log robustness of PSA, p70p, Cyclin D, and AKTp versus protein knock-out. A CR LNCaP cell was assumed for all knock-out cases.}
\label{fg:Dual_Knockout}
\end{figure}

>>>>

We also used robustness analysis to observe the effect of knocking out combinations of key species found in our sensitivity analysis. 
We calculated the effect of perturbation in CR cells on different protein markers in the following 7 cases: (1) Raf knock-out, (2) PI3K knock-out, (3) AR knock-out, (4) Raf and PI3K knock-outs, (5) Raf and AR knock-outs, (6) PI3K and AR knock-outs, and (7) Raf, PI3K and AR knock-outs. Over the 500 ensemble member set, the overall decrease in PSA expression levels was greatest in all cases which included the knock-out of AR (Fig. \ref{fg:Dual_Knockout}A). 
The decreased expression of activated p70  was most prominent in the double PI3K/AR knock-out case and the triple Raf/PI3K/AR knock-out case (Fig. \ref{fg:Dual_Knockout}B). 
Although the change in expression level of cyclin D varied for individual ensemble members (increased and decreased), the median and mean behavior for all knock-out cases is around zero suggesting no change due to perturbation (Fig.\ref{fg:Dual_Knockout}C). 
Individual ensemble members do show a decline in cyclin D expression in cases involving PI3K in double and triple knock-outs and in these cases we also see fewer ensemble members that have an increased expression of cyclin D.  
Fig. \ref{fg:Dual_Knockout}D shows that in the PI3K knock-out case we see a complete inhibition of AKTp expression (a value of -1 is essentially zero expression due to knock-out). 
This is a trivial result in the model because AKTp only occurs through PIP3 recruitment to the membrane, and PIP3 is only activated by PI3K in the model. 
All other knock-out cases show a median response of no fold change in AKTp expression, but we do see a group of ensemble members that have an increased expression of AKTp in the RAF knock-out and RAF/AR knock-out cases. 
Overall, these results are consistent with our hypothesis that a combination therapy will be more effective.

@@

The results of the model suggest that an inhibition of either the PI3K pathway or the MAPK pathway in combination with an AR inhibitor as a possible therapy for CRPC. 
Sensitivity analysis revealed no change in the top sensitive species in the presence or absence of the AR inhibitor, enzalutamide. 
PI3K/AKT and MAPK species continued to fall in the top two percent of sensitive species. 
A study by Carver \textit{et al.} looked at dual inhibition of AR and PI3K signaling in LNCaP cells and in a Pten-deficient murine prostate cancer model \cite{Parker2013}. 
Using both the PI3K inhibitor, BEZ235, and the AR inhibitor, MDV3100 (enzalutamide), the group saw a drastic decrease in the total number of cells. 
Each inhibitor on it’s own had a much smaller effect on total cell number. They saw an increase in the cell death marker, c-PARP, in the dual inhibition case. 
The group hypothesized that AKT inhibition leads to increased AR signaling activity through increased protein concentrations of HER3. 
On the other hand AR inhibition leads to increased AKT activity due to the down regulation of PHLPP, a protein phosphatase that regulates AKT. 
For the simplicity of this model, the HER3 pathway and also cell death were not included in the model. 
Dual knock-out studies of PI3K and AR in our model show only a slight additive effect on cell cycle protein cyclin D through the dual knock-out compared to solely inhibiting PI3K (Fig. \ref{fg:Dual_Knockout}).
The triple case, with PI3K, AR, and RAF knock-outs, also only showed a slight additive effect (Fig. \ref{fg:Dual_Knockout}).
%The pathways in our model appear to be uncoupled and therefore little synergy is shown in the dual knock-out case. 
This could indicate that the combined decrease in cell population due to the dual inhibition of PI3K and AR is entirely due to cell death. 
The Carver \textit{et al.} study did not consider cell cycle proteins or cell growth. 
Our model does show a decrease in cell cycle proteins in the PI3K knock-out as well as in the PI3K and AR dual knock-out case in some ensemble members (Fig. \ref{fg:Dual_Knockout}). 
This result seems to be consistent with the decreased cell count in the PI3K knock-out case which is not dependent on cell death, as c-PARP levels are low. 
The decrease in cell cycle proteins in the model is due to a decrease in general translation, including free eIF4E levels and activated 40S ribosome subunit. 
The decrease in p70 (S6) activation due to inhibition of PI3K is shown in both the model and in the Carver \textit{et al.} study , indicating this result is due to the PI3K pathway (Fig. \ref{fg:Supp_KO}). 
Our experimental results confirm the Carver \textit{et al.} study in that a dual inhibition of AR and PI3K signaling led to a more prominent decrease in cell viability then each of the inhibitors alone in LNCaP cells. 
We extended the study to look at the addition of a third inhibitor, sorafenib, that inhibits Raf kinase in the MAPK pathway and an additional cell line, C4-2, which was CR.  In both cell lines, there was not a significant decrease in cell viability between the three dual inhibitor cases and the triple inhibition case at 74 hours, indicating that a dual inhibition (PI3K/AR, AR/MAPK, or PI3K/MAPK) may be a sufficient treatment.

@@

\begin{figure}\centering
\includegraphics[width=1.0\textwidth]{./figs/Fig_7_BoxPlot_KnockOut_Cases_new.pdf}
\caption{Robustness analysis of a population of CR prostate models with seven protein knock-out cases (N = 500). A log fold change of greater then zero implies that the concentration of the protein increased with the knock-out, while a log fold change of less then zero indicates that the concentration of protein decreased. A log of fold change equal to 0, shows no response due to the knock-out.
 A.,B.,C.,D. Log robustness of PSA, p70p, Cyclin D, and AKTp versus protein knock-out. A CR LNCaP cell was assumed for all knock-out cases.}
\label{fg:Dual_Knockout}
\end{figure}

@@

\begin{figure}\centering
\includegraphics[width=1.0\textwidth]{./figs/Figure8_ExperimentalResults.pdf}
\caption{Experimental results for multiple drug combinations on two prostate cancer cell lines, LNCaP and C4-2. A. Western blot analysis of AR, pS6, AKT, pAKT, ERK and cleaved PARP in LNCaP and C4-2 cell lines treated for 24 hrs with DMSO (control), sorafenib (5 $\mu$M), LY294002 (40 $\mu$M), and MDV3100 (10 $\mu$M) alone or in combination (at least 3 repeats). B. Cells (LNCaP and C4-2) were treated for 24, 48 and 72 hrs with sorafenib (5 $\mu$M), LY294002 (40 $\mu$M), and MDV3100 (10 $\mu$M) and cell viability was measured using MTT Assay. Values were normalized to DMSO (control). C. Cell viability results for LNCaP and C4-2 cells at varying concentration of sorafenib, LY294002, and MDV3100 after 24 hrs of treatment. Values were normalized to DMSO (control). Error bars represent standard error (at least 3 repeats with triplicates performed in each experiment).}
\label{fg:Experiment}
\end{figure}
